Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-12-22 main.tex
begin transformation PH to AAN
continue transformation (encodings + examples + conclusion)
almost finish transformation (except appendix)

% Pretty references
\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{tab}{Table~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{th}{Theorem~\ref{#1}}
\newrefformat{co}{Corollary~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{cr}{Condition~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

@@

% Macros
\newcommand{\sN}{\mathbb{N}}
\newcommand{\sNN}{\mathbb{N}^\bullet}
\def\DEF{\stackrel{\Delta}=}
\def\EQDEF{\stackrel{\Delta}\Leftrightarrow}
\newcommand{\segm}[2]{\llbracket #1; #2 \rrbracket}
\def\indexes#1{\mathbb{I}^{#1}}
\def\f#1{\mathsf{#1}}
\def\prio{\mathsf{prio}}
\newcommand{\bottom}{\perp}
\newcommand{\stable}{\mathsf{stable}}
\newcommand{\update}{\mathsf{update}}
\newcommand{\components}{\Gamma}
%\newcommand{\cs}{CS}
\newcommand{\cs}{\Delta}

@@

% Ajouts et progression
\def\modMF#1{\textcolor{teal}{#1}}
\def\modLP#1{\textcolor{magenta}{#1}}
\def\modMM#1{\textcolor{blue}{#1}}
\def\modOR#1{\textcolor{olive}{#1}}
%\def\modMF#1{#1} \def\modLP#1{#1} \def\modMM#1{#1} \def\modOR#1{#1}

@@

\begin{abstract}
The Process Hitting is a recently introduced framework designed for the modelling of concurrent systems.
Its originality lies in a compact representation of both components of the model and its corresponding actions:
each action can modify the status of a component, and is conditioned by the status of at most one other component.
This allowed to define very efficient static analysis based on local causality to compute reachability properties.
However, in some cases, this modelling lacks some preemption features
that would allow to model relative time constraints between interactions.

In this paper, we propose an extension of the Process Hitting framework
by using classes of priorities in order to improve its expressivity.
Furthermore, we develop a new refinement for the under-approximation of the static analysis
to give accurate results on Process Hitting models.
This method applies to a specific sub-class of models,
but this class turns out to be sufficiently expressive to represent any
Process Hitting model with an arbitrary number of classes of priorities,
thus guaranteeing a wide field of application of our method.

Finally, we also show that the priorities
allow to represent accurate Boolean gates in order to model
cooperations between components
(that is, for instance, when two components are supposed to interact with a third one
only when they are in a given configuration)
which was not possible before.
This allows to accurately model
Asynchronous Discrete Networks, either Boolean or multivalued,
and efficiently under-approximate reachability properties
in these models.
We illustrate this on the model-checking of a signalling network of 94
components, which is unprecedented.
\end{abstract}
\begin{keyword}
discrete networks \sep
abstract interpretation \sep
reachability \sep
qualitative models \sep
systems biology
\end{keyword}
\end{frontmatter}

>>>>

% Pretty references
\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{tab}{Table~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{th}{Theorem~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{co}{Corollary~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{cr}{Condition~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

@@

% Macros
\newcommand{\sN}{\mathbb{N}}
\newcommand{\sNN}{\mathbb{N}^\bullet}
\def\DEF{\stackrel{\Delta}=}
\def\EQDEF{\stackrel{\Delta}\Leftrightarrow}
\newcommand{\segm}[2]{\llbracket #1; #2 \rrbracket}
\def\indexes#1{\mathbb{I}^{#1}}
\def\f#1{\mathsf{#1}}
\def\prio{\mathsf{prio}}
\newcommand{\bottom}{\perp}
\newcommand{\stable}{\mathsf{stable}}
\newcommand{\update}{\mathsf{update}}
\newcommand{\components}{\PHs}
%\newcommand{\cs}{CS}
\newcommand{\cs}{\textcolor{red}{\Delta}}

@@

% Ajouts et progression
\def\modMF#1{\textcolor{teal}{#1}}
\def\moda#1{\textcolor{blue}{#1}}
% \def\modLP#1{\textcolor{magenta}{#1}}
% \def\modMM#1{\textcolor{blue}{#1}}
% \def\modOR#1{\textcolor{olive}{#1}}
%\def\modMF#1{#1} \def\moda#1{#1} \def\modLP#1{#1} \def\modMM#1{#1} \def\modOR#1{#1}

@@

\begin{abstract}
The Process Hitting is a recently introduced framework
designed for the modelling and analysis of concurrent systems.
The particular form of its actions (encoding the dynamics into the models)
allowed to define very efficient static analysis
based on local causality to compute reachability properties.
However, this modelling shows its limits when preemption features are needed,
and especially for cooperations
(that is, the regulation of two components that jointly interact with a third one).

In this paper, we focus on
the Asynchronous Automata Network (AAN) formalism,
which is an extension of the Process Hitting with an increased expressivity.
Like the Process Hitting,
the originality of AANs is the restrictions of the actions
that have an effect limited to exactly one component;
but unlike Process Hitting, their causality is not restricted to at most one other component.

We then develop a new refinement for the under-approximation of the static analysis
to efficiently compute results on the dynamics of AANs.
Because of the expressivity improvements,
this analysis gives more accurate results
than the previous works carried on the Process Hitting.
Furthermore, its complexity is still polynomial in the size of the model
and exponential in the number of possible “branchings” during the solving.

We also extend the applicability of this static analysis to
AANs with classes of priorities and to
Asynchronous Discrete Networks, either Boolean or multivalued,
which allows to efficiently study the dynamics of these widespread models
% efficiently under-approximate reachability properties
% in these widespread models.
We illustrate this on the model-checking of a signalling network of 94
components, which is unprecedented.
\end{abstract}
\begin{keyword}
discrete networks \sep
abstract interpretation \sep
reachability \sep
qualitative models \sep
systems biology
\end{keyword}
\end{frontmatter}

### 2014-12-22 parts/def-ph.tex
almost finish transformation (except appendix)

We give in this section the definition and the semantics of the Asynchronous Automata Networks (AAN).
It is a restriction of the classical Synchronous Automata Networks
where each set of transitions sharing the same label
can only change one local state at a time.
\todo{We also discuss how it is related to the local state Hitting framework (with priorities)
\cite{PMR10-TCSB,FPMR13-CS2Bio}.}
As shown in \pref{sec:encodings},
this class of models is equivalent to Asynchronous Discrete Networks,
and extending it with classes of priorities does not increase its expressivity.

@@

>>>>

We give in this section the definition and the semantics of the Asynchronous Automata Networks (AAN).
It is a restriction of the classical Synchronous Automata Networks
where each set of transitions sharing the same label
can only change one local state at a time.
We also discuss how it is related to the Process Hitting framework
(with or without classes of priorities).
In \pref{sec:encodings}, we will also introduce classes of priorities into these models,
and show that they are equivalent to Asynchronous Discrete Networks.

@@

We note that obviously, Process Hitting~\cite{PMR10-TCSB} being a restriction of AANs,
one can easily encode any Process Hitting model as an AAN by using only actions
$\PHfrappe{A}{b_j}{b_k}$ so that $0 \leq \card{A} \leq 1$.
Process Hitting with classes of priorities~\cite{FPMR13-CS2Bio} can also be encoded as AANs
with priorities, as defined in \pref{sec:encodings},
by using the same priorities for each action,
and possibly translate them into simple AANs.

### 2014-12-22 parts/intro.tex
almost finish transformation (except appendix)

Discrete modelling frameworks for biological networks is an active research field where formal methods have proven that they were very powerful.
Such a work started in the seventies.
It was later enriched in many directions and widely used to elucidate many biological questions.
Among these questions, a major one is to understand precisely how biological systems evolve and behave; why and how they change their usual behaviours, etc.
This leads to questions about the (possible or inevitable) reachability of some states.
The ultimate goal is to discover how it could be possible to prevent biological systems from reaching some pathological states.
\todo{Ajouter des références}

Of course, such formal models on which analyses are performed are abstract representations of the actual studied systems.
They are associated with parameters that have to be synthesised %so as to be as much as possible fitting with the real systems having some observed behaviors.
to give the most faithful representation of the real systems with their observed behaviours.
As a matter of fact, the abstractions we get are more or less rough or accurate.
Prevalent formal frameworks for such modelling activities are state-transition systems or process algebras. % Petri nets
We developed a quite similar framework named the Process Hitting \cite{PMR10-TCSB},
consisting in a restriction of these frameworks where the evolution of a component is determined by the state of at most one other component that does not evolve.
In a sense, these kind of actions are of the form $X + Y \rightarrow X + Y'$ where $X$ behaves like a catalyst molecule that “hits” another molecule $Y$ and changes it into $Y'$, without being changed itself.
Assuming catalysts are always available, this can represent any biochemical system made of monomolecular reactions, and can also represent catalytic networks such as metabolic networks.
%, with the aim of avoiding to build the whole state space in order to be able to tackle very large systems
Our motivation behind this framework was to design a model and analysis techniques adapted to biological modelling.
These analyses avoid to build the whole state space, which allows to tackle very large systems (that would have led to a huge number of states, hopelessly too huge to be analysed).
They are based on the fact that most biological models have few levels of expression per component:
in Boolean networks~\cite{kauffman69,Thomas73} there are only two levels per component, and in their multivalued equivalent~\cite{deJong02}, components rarely have more than four levels.

Besides, one further objective of our work is now %to be more accurate in the description of the studied systems.
to improve the accuracy of the description of the studied systems dynamics.
The idea for this is to introduce timing features into models:
we are interested in taking into account some knowledge about the relative length of some phenomena as it is a way to refute some models (or parameters) that are inconsistent with the observed dynamic behaviours.
In this paper, we are dealing with these timing properties through priorities,
that are based on the simple founding idea that actions with higher priority have to be processed before the ones with lower priority.
Furthermore, due to the Process Hitting framework restrictions, multimolecular reactions are not immediately available, but one can simulate them with an encoding called “cooperation”.
That encoding however introduces extra reactions, and this is where the priorities become useful, if not necessary.
The extra reactions can be given “infinite speed” (highest priority) so that they do not affect the behaviour of “normal” (lower priority) reactions, including the multimolecular ones.

Until now, such a priority scheduling of the actions was not studied extensively in the different formal modelling frameworks dedicated to systems biology.
Nevertheless, such an attempt has been carried out for Petri nets by F.~Bause \cite{Bause97},
and the concept of priority relations among the transitions of a network has also more recently been introduced by A.~K.~Wagler \textit{et~al.} \cite{waw,WaglerW12} in order to allow modelling deterministic systems for biological applications.
The concept of priority is much straightforward in the approach of process algebras as it was shown by R.~Cleaveland and M.~Hennessy in \cite{Cleaveland199058,Cleaveland99prioritiesin} and their abstractions and equivalences were studied in \cite{Cleaveland:2007:PAP:1282576.1282847}.
It was later extended for applications in the field of systems biology by M.~John \textit{et~al.} \cite{jlnu2010}.

\subsection*{Contributions}
\todo{Compléter les contributions}
Since our formalism (the Process Hitting) can be considered as a subset of Calculus of Communicating Systems, %is inspired from the $\pi$-calculus,
our work is related to such semantic ramifications of extending traditional process algebras with the concept of priority that allows for some transitions to be given precedence over others.
The concept is derived in two directions: dynamic versus static, the difference being naturally that the former one refers to a semantics where priority values may change during execution according to some evolution rules.
In our work, actions exhibit a static priority structure.
The actions modelling biological processes are divided in several classes of priorities which stand for the different reaction rates or relative preemptions.
Furthermore, the highest class of priority is used to model particular nonbiological actions, designated as “prioritised”,
that are used to encode Boolean functions inside the model.

This paper is an extension of \cite{FPMR13-CS2Bio} where we
introduced a new extension to the semantics of Process Hitting by partitioning actions into classes of priorities
(\pref{sec:ph}).
One of the objectives is to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics, using only two classes of priorities.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability of the state of components on a subclass of this new framework, thus allowing to compute efficient static analysis.
This local reachability under-approximation can be also easily extended to study the reachability of a global state.
Furthermore, we provide a new extended under-approximation method to check a successive reachability property.

Whereas the analysis presented in \pref{sec:sa} considers a particular class of
Process Hitting with two classes of priority,
we provide in \pref{sec:encodings} a “flattening” of any Process Hitting model into
this class of models,
so that its application is not restricted to it.
We also give in the same section a formal encoding of Asynchronous Discrete Networks,
that is, the multivalued version of Asynchronous Boolean Networks.
These models,
which encompass variables and evolution functions with a limited number of discrete values,
also called “Logical Networks” in~\cite{Thomas95,deJong02}.
This encoding thus allows to efficiently study this kind of widespread models.

In \pref{sec:example}, after giving an example of the encoding of a biological
network with multiple priority classes,
we apply our method on a large-scale biological model containing 94 components.
The under-approximation turned out to be conclusive in all cases and results were computed in hundredths of seconds,
thus overtaking the efficiency of usual model-checkers.

\modMF{
The main addition in this paper compared to \cite{FPMR13-CS2Bio} are the following:
\pref{th:bisimPHP} states that any Process Hitting model with any number of classes of priorities
can be represented with only two classes of priorities,
thus extending the scope of our method to any possible Process Hitting model;
\pref{thm:ordered-ua} gives a method to cope with successive reachability,
and obtain more conclusive results in the case of several successive objectives;
\todo{def extraction} allows to extract a concrete scenario when our static analysis method
is conclusive.
%All aim at extending the application range of the static analysis we developed.
}

>>>>

Discrete modelling frameworks for biological networks is an active research field where formal methods have proven that they were very powerful.
Such a work started in the seventies.
It was later enriched in many directions and widely used to elucidate many biological questions.
Among these questions, a major one is to understand precisely how biological systems evolve and behave; why and how they change their usual behaviours, etc.
These questions are strongly linked to the (possible or inevitable) reachability of some states.
The ultimate goal is to discover how it could be possible to prevent biological systems from reaching some pathological conditions.
\todo{Ajouter des références}

Of course, such formal models on which analyses are performed are abstract representations of the actual studied systems.
They are associated with parameters that have to be synthesised %so as to be as much as possible fitting with the real systems having some observed behaviors.
to give the most faithful representation of the real systems with their observed behaviours.
As a matter of fact, the abstractions we get are more or less rough or accurate.
Prevalent formal frameworks for such modelling activities are state-transition systems or process algebras. % Petri nets
We developed a quite similar framework named the Process Hitting~\cite{PMR10-TCSB},
consisting in a restriction of these frameworks where the evolution of a component is determined by the state of at most one other component that does not evolve.
In a sense, these kind of actions are of the form $X + Y \rightarrow X + Y'$ where $X$ behaves like a catalyst molecule that “hits” another molecule $Y$ and changes it into $Y'$, without being changed itself.
Assuming catalysts are always available, this can represent any biochemical system made of monomolecular reactions, and can also represent catalytic networks such as metabolic networks.
%, with the aim of avoiding to build the whole state space in order to be able to tackle very large systems
Our motivation behind this framework was to design a model and analysis techniques adapted to biological modelling.
These analyses avoid to build the whole state space, which allows to tackle very large systems (that would have led to a huge number of states, hopelessly too huge to be analysed).
They are based on the fact that most biological models have few levels of expression per component:
in Boolean networks~\cite{kauffman69,Thomas73} there are only two levels per component, and in their multivalued equivalent~\cite{deJong02}, components rarely have more than four levels.

Besides, one further objective of our work is now %to be more accurate in the description of the studied systems.
to improve the accuracy of the description of the studied systems dynamics.
The idea for this is to introduce timing features into models:
we are interested in taking into account some knowledge about the relative length of some phenomena as it is a way to refute some models (or parameters) that are inconsistent with the observed dynamic behaviours.
In this paper, we are dealing with these timing properties through priorities,
that are based on the simple founding idea that actions with higher priority have to be processed before the ones with lower priority.
Furthermore, due to the Process Hitting framework restrictions, multimolecular reactions were previously not immediately available, but one could simulate them with an encoding called “cooperative sort”.
That encoding however introduces extra reactions,
that produce a temporal shift between the presence of the reactants
and the playability of the reaction.
This is where the priorities become useful, if not necessary:
the extra reactions can for example be given “infinite speed” (highest priority) so that they do not affect the behaviour of “normal” (lower priority) reactions, including the multimolecular ones.
Another approach consists in considering another class of models,
that we call Asynchronous Automata Network,
and that allows to naturally model these cooperations
by defining several requisites for a reaction.
Furthermore, such automata networks are still compatible with the notion of priority,
that can also be used to model different reaction rates in the model.

Until now, such a priority scheduling of the actions was not studied extensively in the different formal modelling frameworks dedicated to systems biology.
Nevertheless, such an attempt has been carried out for Petri nets by F.~Bause~\cite{Bause97},
and the concept of priority relations among the transitions of a network has also more recently been introduced by A.~K.~Wagler \textit{et~al.}~\cite{waw,WaglerW12} in order to allow modelling deterministic systems for biological applications.
The concept of priority is much straightforward in the approach of process algebras as it was shown by R.~Cleaveland and M.~Hennessy in~\cite{Cleaveland199058,Cleaveland99prioritiesin} and their abstractions and equivalences were studied in~\cite{Cleaveland:2007:PAP:1282576.1282847}.
It was later extended for applications in the field of systems biology by M.~John \textit{et~al.}~\cite{jlnu2010}.

\subsection*{Contributions}
Since our formalisms (the Asynchronous Automata Network and the Process Hitting)
can be considered as a subset of Calculus of Communicating Systems,
or Synchronous Automata Networks,
%is inspired from the $\pi$-calculus,
our work is related to such semantic ramifications of extending traditional process algebras with the concept of priority that allows for some transitions to be given precedence over others.
The concept is derived in two directions: dynamic versus static, the difference being naturally that the former one refers to a semantics where priority values may change during execution according to some evolution rules.
In our work, actions exhibit a static priority structure.
The actions modelling biological processes are divided in several classes of priorities which stand for the different reaction rates or relative preemptions.
Furthermore, the multiple causality of the actions in Asynchronous Automata Networks
allow to abstract a highest class of priority that would be used
to model particular nonbiological actions
that encode Boolean functions inside the model.

This paper is an extension of~\cite{FPMR13-CS2Bio} where we
introduced a new extension to the semantics of Process Hitting
by partitioning actions into classes of priorities.
One of our objectives was to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics.
Although in~\cite{FPMR13-CS2Bio} this was achieved with a particular class of high priority,
the work presented in this paper, however, focuses on Asynchronous Automata Networks.
The simple version of this formalism (without classes of priorities)
is defined in \pref{sec:ph}
and consists of an extension of Process Hitting with classes of priorities
that allows to simplify the formalisation with the same benefits.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability
of several local states of components,
thus allowing to compute efficient static analysis of the dynamics.
This local reachability under-approximation can be also easily extended to study the reachability of a global or partial state.
Furthermore, we provide a new extended under-approximation method to refine this method
in the case of a successive reachability property.

Furthermore,
we define in \pref{sec:encodings} the notion of Asynchronous Automata Networks
with classes of priorities,
that consists of a convenient way to introduce time parameters
into the models, under the form of preemptions.
We also provide in \pref{sec:encodings} a “flattening”
of any Asynchronous Automata Networks with classes of priorities
into a model without classes of priorities,
thus extending the range of applicability of the static analysis.
We also give in the same section a formal encoding of Asynchronous Discrete Networks,
that is, the multivalued version of Asynchronous Boolean Networks.
These models,
which encompass variables and evolution functions with a limited number of discrete values,
are also called “Logical Networks” in~\cite{Thomas95,deJong02}.
This encoding thus allows to efficiently study this kind of widespread models.

In \pref{sec:example}, we give two examples of the use of our method.
The first consists of a detailed application of the analysis of a biological network expressed as
an ANN with 2 classes of priorities,
which requires to be encoded into a simple ANN (without priorities).
The second consists of applying our method
on a large-scale biological model containing 94 components;
the under-approximation turned out to be conclusive in all cases and results were computed in hundredths of seconds,
thus overtaking the efficiency of usual model-checkers.

\modMF{
The main addition in this paper compared to~\cite{FPMR13-CS2Bio} is
the representation under the form of ANNs, that allow to simplify the formalisation.
Furthermore, we give new tools to improve the applicability of our method:
\pref{th:bisimPHP} states that any ANN with any number of classes of priorities
can be represented into a simple ANN (or, equivalently, with only one class of priority),
thus extending the scope of our method to any possible Process Hitting model;
\pref{thm:ordered-ua} gives a method to refine successive reachability,
and obtain more conclusive results in the case of several successive objectives;
\pref{def:concret} allows to extract a concrete scenario when our static analysis method
is conclusive.
}

### 2014-12-22 parts/models-ph-dn.tex
almost finish transformation (except appendix)

$f^i_\varsigma$ hits the local states of automaton $a^i$ to make them bounce to the local state
$a^i_{k'}$ if and only if $k'=f^i(\decode \varsigma)$;
$\decode \varsigma$ being the ADN state corresponding to the PH (partial) state $\varsigma$ (note that
$f^i(\decode \varsigma)$ is fully defined because $\decode \varsigma$ specifies the state for all
the components in $\DNdep(f^i)$).

% 
% between the components $\DNdep(f^i)$.
% 
% Two classes of actions are then defined:
% $\Hits^{(1)}$ is the set of actions updating the cooperative automata according to the current state of the
% components:
% if $j\in\DNdep(f^i)$, $a^j_k$ hits each local state $f^i_\varsigma$ where $\get{\varsigma}{a^j}\neq
% a^j_k$ to make it bounce to the local state $f^i_{\varsigma \Cap a^j_k}$.
% $\Hits^{(2)}$ is the set of actions encoding the transitions in the ADN:
% $f^i_\varsigma$ hits the local states of automaton $a^i$ to make them bounce to the local state
% $a^i_{k'}$ if and only if $k'=f^i(\decode \varsigma)$;
% $\decode \varsigma$ being the ADN state corresponding to the PH (partial) state $\varsigma$ (note that
% $f^i(\decode \varsigma)$ is fully defined because $\decode \varsigma$ specifies the state for all
% the components in $\DNdep(f^i)$).

\begin{definition}
\label{def:DN2PH}
  $\toPH\DNdef=(\Sigma,\PHl,\Hits)$ is the AAN encoding the ADN $\DNdef$, with:
  \begin{itemize}
    \item $\Sigma = \{ a^1, \dots, a^n \}$;
%      the automata representing components;

@@

    \item $\PHh = \{ \PHfrappe{A}{a^i_j}{a^i_k} \mid
      \exists \varsigma \PHsubl[\PHl]_{\DNdep(f^i)}, \exists x \in \mathbb{F},
      \varsigma \subseteq \encode{x} \wedge
      f^i(x) = k \wedge \PHget{x}{i} = j \wedge j \neq k \wedge
      A = \toset{\varsigma} \}$ where $\encode{x}$ is defined below.

%     \item $\Hits^{(1)}= \{ \hit{a^j_k}{f^i_\varsigma}{f^i_{\varsigma'}}
%       \mid i\in\segm{1}{n} \wedge
%       j\in\DNdep(f^i) \wedge a^j_k\in \PHl_{a^j} \wedge f^i_\varsigma\in \PHl_{f^i}
%       \wedge
%       \get{\varsigma}{a^j}\neq a^j_k 
%       \wedge \get{\varsigma'}{a^j}=a^j_k
%       \wedge (\get{\varsigma'}{a^{l}}=\get{\varsigma}{a^l},
%       \forall l\in\segm{1}{n},l\neq j)
%       \}$, 
%       the set of actions with priority $1$ that update cooperative automata;
% 
%     \item $\Hits^{(2)}=\{ \hit{f^i_\varsigma}{a^i_k}{a^i_{k'}} \mid i\in\segm{1}{n} \wedge
%           f^i_\varsigma\in \PHl_{f^i} \wedge
%           a^i_k\in \PHl_{a^i}\wedge k\neq k' \wedge f^i(\decode \varsigma) = k'
%         \}$,
%       the set of actions with priority $2$ for updating the components using their respective discrete
%       maps.
%       $\decode \varsigma$ is defined below.
  \end{itemize}
  Given a state $s \in \PHl$ of the AAN,
  $\decode s=x$ is the corresponding state in the ADN:
  $\forall i\in\segm{1}{n}, \get{s}{a^i}=a^i_k \Rightarrow \get{x}{i}=k$.
  Conversely, given a state $x\in \mathbb F$ of the ADN, 
  $\encode x=s$ is the corresponding state in the AAN:
  $\forall i\in\segm{1}{n}, \get{x}{i}=k \Rightarrow \get{s}{a^i}=a^i_k$.
\end{definition}

Finally,
\pref{th:bisimDN} states the weak bisimulation relation between an ADN and its encoding in AAN.
Intuitively, the actions follow strictly the possible transitions of the ADN.
A detailed proof is given in \pref{suppl:proofbisimADN}.

\begin{theorem}[$\DNdef \approx \toPH\DNdef$]~
\label{th:bisimDN}
  \begin{enumerate}
    \item \label{adn2ph} $\forall x,x'\in\mathbb F$,
      $x\DNtrans x' \Longrightarrow \encode x \PHPtrans^* \encode{x'}$,
      where $\PHPtrans^*$ is a finite sequence of $\PHPtrans$ transitions.

    \item \label{ph2adn} $\forall s,s'\in \PHl$,
      $s\PHPtrans s' \Longrightarrow \decode s = \decode {s'} \vee \decode s \DNtrans
      \decode{s'}\enspace.$
  \end{enumerate}
\end{theorem}

>>>>

\begin{definition}[Encoding ($\toPH$)]
\label{def:DN2PH}
  $\toPH\DNdef=(\Sigma,\PHl,\Hits)$ is the AAN encoding the ADN $\DNdef$, with:
  \begin{itemize}
    \item $\Sigma = \{ a^1, \dots, a^n \}$;

@@

    \item $\PHh = \{ \PHfrappe{A}{a^i_j}{a^i_k} \mid
      \exists \varsigma \PHsubl[\PHl]_{\DNdep(f^i)}, \exists x \in \mathbb{F},
      \varsigma \subseteq \encode{x} \wedge
      f^i(x) = k \wedge \PHget{x}{i} = j \wedge j \neq k \wedge
      A = \toset{\varsigma} \}$ where $\encode{x}$ is defined below.
  \end{itemize}
  Given a state $s \in \PHl$ of the AAN,
  $\decode s=x$ is the corresponding state in the ADN:
  $\forall i\in\segm{1}{n}, \get{s}{a^i}=a^i_k \Rightarrow \get{x}{i}=k$.
  Conversely, given a state $x\in \mathbb F$ of the ADN, 
  $\encode x=s$ is the corresponding state in the AAN:
  $\forall i\in\segm{1}{n}, \get{x}{i}=k \Rightarrow \get{s}{a^i}=a^i_k$.
\end{definition}

Finally,
\pref{th:bisimDN} states the weak bisimulation relation between an ADN and its encoding in AAN.
Intuitively, the actions follow strictly the possible transitions of the ADN.
A detailed proof is given in \pref{suppl:proofbisimADN}. \todo{Utile ?}

\begin{theorem}[$\DNdef \approx \toPH\DNdef$]
\label{th:bisimDN}
  If $\DNdef$ is an ADN
  and $\PH = \toPH\DNdef = (\PHs; \PHl; \PHa)$ is its AAN encoding, then:
  \begin{enumerate}
    \item \label{adn2ph} $\forall x,x'\in\mathbb F$,
      $x\DNtrans x' \Longrightarrow \encode x \PHPtrans \encode{x'}$,
    \item \label{ph2adn} $\forall s,s'\in \PHl$,
      $s\PHPtrans s' \Longrightarrow \decode s \DNtrans
      \decode{s'}\enspace.$
  \end{enumerate}
\end{theorem}

### 2014-12-22 parts/sa.tex
almost finish transformation (except appendix)

\modMF{%
First, $\osres$ is used to split the objective sequence $\w$ to solve,
by considering successive objectives, in the fashion of \pref{thm:ordered-ua}.
This function calls $\res$ which allows to solve an objective $P$
by either re-target it (if $\gCont(\PHsort(P), P) \neq \emptyset$)
of by searching for a solution to this objective in $\BS(P)$.
This search may require to try several objective sequences $\zeta \in \BS(P)$.
For each trial, the chosen objective sequence $\zeta$ is solved with $\rec$,
which recursively created a scenario to play the whole sequence.
This scenario is itself recursively created by combining
smaller scenarios allowing to reach each required hitter of each action
with the function $\thisseq$.
Finally, $\thisseq$ calls $\res$ anew to reach the hitters
that are necessarily contained in other automata.
}%

\begin{definition}[$\osres : \PHl \times \OS \to \Sce$]
\modMF{%
For all $s \in \PHl$ and $\w \in \OS$, we define:
\[
  \osres_s(\w) \DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\w = \emptyseq$} \\
      \res_s(\w_1) \concat
      \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
        & \text{ otherwise} \\
    \end{cases}
\]
and, for all $s \in \PHl$ and $P \in \Obj$:
\[\res_s(P) \DEF
  \begin{cases}
    \varepsilon
      & \text{if } \target{P} = \bounce{P} \\
    \osres_s(P^Q \concat Q)
      & \text{if } \exists Q \in \Obj, (P, Q) \in \Bee{\Obj}{\Obj} \\
    \rec_s(\zeta)
      & \text{otherwise, with } \zeta \in \BS(P) \text{\todo{Selon possbilités}}
  \end{cases}
\]
with $P^Q = \PHobj{\PHtarget(P)}{\PHtarget(Q)}$.
Furthermore:
\begin{align*}
  \forall \zeta \in \BS,
  \rec_s(\zeta) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\zeta = \emptyseq$} \\
      \thisseq_s(\zeta_1) \concat \zeta_1 \concat
      \rec_{s \play \thisseq_s(\zeta_1) \concat \zeta_1}
      (\zeta_2 \concat \dots \concat \zeta_{\card{\zeta}})
        & \text{ otherwise}
    \end{cases} \\
  \forall h \in \PHh,
  \thisseq_s(h) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\hitter{h} = \emptyset$} \\
      \R_s(h_1) \concat \thisseq_{s \play \R_s(h_1)}(h_2 \concat \dots \concat h_{\card{h}})
        & \text{ otherwise}
    \end{cases} \\
  \forall a_i \in \Proc,
  \R_s(a_i) &\DEF \res_s(\PHobj{\PHget{s}{\PHsort(a_i)}}{a_i})
\end{align*}
and for all action $h \in \PHh$, $(h_i)_{i \in \segm{1}{\card{\hitter{h}}}}$
is a sequence containing all hitters of $h$ in an arbitrary order.
}
\end{definition}

\todo{Gérer les choix multiples possibles de $\zeta \in \BS(P)$}

>>>>

\modMF{%
First, $\osres$ is used to split the objective sequence $\w$ to solve,
by considering successive objectives, in the fashion of \pref{thm:ordered-ua}.
This function calls $\res$ which allows to solve an objective $P$
by either re-target it (if $\gCont(\PHsort(P), P) \neq \emptyset$)
of by searching for a solution to this objective in $\BS(P)$.
This search may require to try several objective sequences $\zeta \in \BS(P)$.
For each trial, the chosen objective sequence $\zeta$ is solved with $\rec$,
which recursively created a scenario to play the whole sequence.
This scenario is itself recursively created by combining
smaller scenarios allowing to reach each required hitter of each action
with the function $\thisseq$.
Finally, $\thisseq$ calls $\res$ anew to reach the hitters
that are necessarily contained in other automata.
In the end, a scenario can be obtained by analysing all values of
$\osres_s(\w)$ (\pref{def:concret})
for all possible initial state $s \subseteq \ctx$ in the considered context,
and all possible bounce sequences $\zeta \in \BS(P)$ that are chosen during the resolution.
}%

\begin{definition}[$\osres : \PHl \times \OS \to \Sce$]
\label{def:concret}
\modMF{%
For all $s \in \PHl$ and $\w \in \OS$, we define:
\[
  \osres_s(\w) \DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\w = \emptyseq$} \\
      \res_s(\w_1) \concat
      \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
        & \text{ otherwise} \\
    \end{cases}
\]
and, for all $s \in \PHl$ and $P \in \Obj$:
\[\res_s(P) \DEF
  \begin{cases}
    \varepsilon
      & \text{if } \target{P} = \bounce{P} \\
    \osres_s(P^Q \concat Q)
      & \text{if } \exists Q \in \Obj, (P, Q) \in \Bee{\Obj}{\Obj} \\
    \rec_s(\zeta)
      & \text{otherwise, with } \zeta \in \BS(P)
  \end{cases}
\]
with $P^Q = \PHobj{\PHtarget(P)}{\PHtarget(Q)}$.
Furthermore:
\begin{align*}
  \forall \zeta \in \BS,
  \rec_s(\zeta) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\zeta = \emptyseq$} \\
      \thisseq_s(\zeta_1) \concat \zeta_1 \concat
      \rec_{s \play \thisseq_s(\zeta_1) \concat \zeta_1}
      (\zeta_2 \concat \dots \concat \zeta_{\card{\zeta}})
        & \text{ otherwise}
    \end{cases} \\
  \forall h \in \PHh,
  \thisseq_s(h) &\DEF
    \begin{cases}
      \emptyseq
        & \text{ if $\hitter{h} = \emptyset$} \\
      \R_s(h_1) \concat \thisseq_{s \play \R_s(h_1)}(h_2 \concat \dots \concat h_{\card{h}})
        & \text{ otherwise}
    \end{cases} \\
  \forall a_i \in \Proc,
  \R_s(a_i) &\DEF \res_s(\PHobj{\PHget{s}{\PHsort(a_i)}}{a_i})
\end{align*}
and for all action $h \in \PHh$, $(h_i)_{i \in \segm{1}{\card{\hitter{h}}}}$
is a sequence containing all hitters of $h$ in an arbitrary order.
}
\end{definition}
